Integrated approach to modeling krill population dynamics in the Western Antarctic Peninsula: spatial and ecosystem considerations

### 2024-09-11 SS3_Krill_481_FSA.Rmd
finish edits (i hope)

---
title: "Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
author:
  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Milenio Base]
  - Lucas Krüger ^[Instituto Antártico Chileno, Milenio Base]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Milenio Base]
  - George Watters^[Ecosystem Antarctic Division, Southwest Fisheries Science Center, NOAA, USA]

date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
always_allow_html: true
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

@@

The integration of available sources of information is a challenge for any attempt to know dynamics populations for Antarctic Krill *Euphausia superba* in an integrated stock assessment model.Being able to identify sources of information, systematize and integrate it in a model and estimate krill population variables in a context of change and fisheries management is vital for management and decision making. This approach has three main objectives. Firstly, integrate fishing, environmental and ecological variables in an integrated stock assessment model and test his performance. On the other hand, consider the spatial heterogeneity of the krill population structure. Finally, to evaluate the impact of biological and structure population assumptions on the performance of this type of integrated stock assessment models. Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region. This study addresses key recommendations highlighted in the preliminary report of the Working Group on Statistics, Assessment, and Modelling (WG-SAM-2024/26), where main observations were: 1) Concerns about using a length-based model due to the scarcity of direct age data, as this approach avoids the approximations required when converting length data to age data; 2) The need to consider the impact of changes in trawl designs over time and to reassess the assumption that the Antarctic Peninsula is a closed system; 3) The limitations of using krill length and biological composition data from fishing vessels, given the significant variability and potential biases, and the importance of understanding the interaction between fisheries and krill-dependent predators; 4) A recommendation to present a progression of model implementations from simple to complex to better evaluate model assumptions, investigate patterns in residuals, and assess the realism of parameter values used, while also integrating recent findings on krill stock hypotheses, with special attention to harvest rate and management procedures. Each observation will be explained in theoretical terms and, where possible, analyzed in detail (Appendix). This approach reflects an ongoing work in progress that is continuously refined based on interactions with the CCAMLR Working Groups. These interactions provide valuable feedback that helps to shape and improve the study's methodology and findings. It should be noted, of this approach is vital for contributing to management strategies based on quantitative methods that provide indicators of krill health, integrating ecosystem aspects into the assessment.

*Keywords: Krill populations, dynamic population, integrated model, Stock Synthesis (SS3), spatial implicit, ecosystem consideration, management, CCAMLR.*

@@

Antarctic krill (*Euphausia superba* krill thereafter) population is predominantly concentrated in the Western Antarctic Peninsula (WAP), serving as a vital food source for predators and being targeted by increasing fishing activities. On the other hand, WAP is considered one of the most sensitive areas to Climate Change (CC) and has already experienced fast changes in various dimensions [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a]. Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities. As part of its mandate, CCAMLR promotes sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013]. CCAMLR has tried to conduct different modeling approaches to assess the status of important species, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016]. Various stock assessment approaches on krill have been conducted to provide management advice to CCAMLR using model-based methodologies. Many of these models face challenges in aligning with management schemes or accurately representing the population dynamics of krill, due to its variability and heterogeneity in spatial distribution and ecosystem complexity. Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys [@Green2023; @Watters2013]. However, these approaches may overlook critical aspects of krill dynamics, such as integrated different data components, spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions. There is a growing consensus on the need for using stock assessment models that incorporate multiple sources of information, capture spatial heterogeneity and consider the complex interplay of ecosystem variables in krill. Integrating the complexity of krill population dynamics presents several challenges. Firstly, data consistency over the years has been problematic, as many monitoring programs and studies have not adhered to a uniform standard throughout their temporal and spatial extension. Additionally, certain regions or areas lack sufficient data or possess only limited information regarding krill dynamics. Furthermore, interpreting indirect indicators that may affect the krill population, such as environmental variables, predator abundance, and fishery data, adds another layer of complexity. Despite these gaps, the existing data about krill population dynamics is abundant, considered data rich, and this availability allows building an integrated model to get population signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others. In response to this need, we propose a conceptual model to conduct a statistical model in Antarctic krill population in the Antarctic Peninsula, which in turn allows the application of an integrated model that incorporates the fishery, surveys and ecosystem information on krill population over the past 40 years.

@@

The study area for the assessment of krill populations in the Antarctic Peninsula is Subarea 48.1 and strata subdivision to account for spatial heterogeneity [@Dornam2021] (Figure \@ref(fig:mapa)). The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill. Therefore, dividing the study area into strata allows to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.

```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
knitr::include_graphics('Figs/481.png')
```

@@

Multiple sources of data to inform the population dynamics model are included in this analysis. These data sources include those listed in Table \@ref(tab:datasources). These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program. This database has information about vessel, nationality, georeferenced, among others; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including: data on abundance by strata, and biological characteristics of krill collected through net sampling. All this data was processed in @Dornam2021; and length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl. Mat. 2; 3). For this analysis, chlorophyll-a concentration (Chl) was assessed as a primary environmental driver. Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U. Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148). This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and  krill dynamics in the WAP. We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + Sea Ice Concentration (SIC) + Sea Surface Temperature (SST)). The model included spatial component (cell) as random effect (formula: ~1 | cell). Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020] (Suppl. Mat 3); 4) Predator abundance data from "MAPPPD". Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.

@@

Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2). Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].

@@

A conceptual model for krill in SubArea 48.1 includes the five spatial strata described previously. Despite differences among strata, the krill population is closed for emigration and immigration, and their spatial distribution and interactions occur within this single, cohesive population. There are two important issues related to the current stock assessment approach. The first one, while we understand that there are krill flows to and from Area 48.1, the challenge of having time series data to implement a model that covers most of Subarea 48 becomes difficult. However, we note that as monitoring programs provide the necessary data to implement models of this nature, we are open to a more comprehensive approach in the Southern Ocean. Secondly, the current model implemented in `SS3` considered the spatial heterogeneity of the krill stock structure within Area 48.1 [@WGEMM2439], however, a model that involves movement between strata has not yet been developed, and is part of the ongoing work presented at the @CCAMLRWGSAM2024. Ecological and environmental variables that have an impact on the krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels. The model considers interannual fluctuations of growth dynamics , maturity and reproduction, with a spawning pulse per year.

@@

Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables. The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21. SS3 is a widely used software tool for assessing fish and invertebrate populations widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions. SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data. The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters. The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population. All analysis were executed in  R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019]  and *ss3diags* packages [@Winker2023]. Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit).

\newpage 
```{r parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}

@@

parbio<-ctl1$MG_parms[1:10,c(1:4,7)]
 row.names( parbio)<-c("Nat M",
                       "Lmin", 
                       "Lmax",
                       "VonBert K",
                       "CV young",
                       "CV old", 
                       "Wt a", 
                       "Wt b",
                       "L50%", 
                       "Mat slope")

 SRpar<-ctl1$SR_parms[1:5,c(1:4,7)]
 Qpar<-ctl1$Q_parms[1:2,c(1:4,7)]
 Selpar<-ctl1$size_selex_parms[1:22,c(1:4,7)]
 parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))

@@

Like most invertebrates, krill is a species for which aging is complex [@Punt2013]. Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure. Regarding this, we consider that a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data in krill monitoring. Aging techniques for krill have not been available historically, although some recent progress has been made [@Kilada2017]. In this sense, when reading of hard parts is difficult like in krill, catch-at-length data are more plentiful, as the collection of length information is relatively cheap. Length offers insights into the age structure of the population, as there is a correlation between age and length.  So far, all integrated models for krill have been of the Catch-at-Length type [@Kinzey2015a; @Kinzey2019a; @Wang2021]. This means they are approaches that internally model age through growth parameters and the Age-length-key (ALK). To perform this conversion, various methods have been used, but the AKL matrix is perhaps the most important one [@ICCA2003]. ALK are generated by aging a sub-sample of a population and used to estimate the age distribution from larger length samples. ALKs assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases. Seasonal or multi-annual applications require careful justification, as using an ALK from a single period can introduce significant bias. Length-stratified sampling is essential, and the process of developing ALKs is labor-intensive, requiring optimal data collection to ensure accuracy  [@ICCA2003]. Formulae exist to estimate the number of age determinations and length measurements is calculated as;

@@

```{r AKL, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in 48.1"}
knitr::include_graphics('Figs/AKL.png')
```

@@

Since we don't have age data monitoring program, we plan to explore *platoon* (Figure \@ref(fig:platoon)) alternative to compare its performance and results with the current length-based model using currently in WG-SAM-2024/26.

```{r platoon, out.width='70%', fig.cap="Reprsentation of platoons in Krill lenght frencuecies"}
x_values <- seq(0, 7, by=0.1)
means <- c( 2.9,  3.5, 4.1)
std_devs <- c(0.65, 0.8, 0.65)
data <- data.frame(
  x = rep(x_values, each = length(means)),
  mean = rep(means, times = length(x_values)),
  std_dev = rep(std_devs, times = length(x_values))
)

@@

According to the observations made at @CCAMLRWGSAM2024 regarding changes in net size and trawl type that could affect selectivity, we have incorporated a scenario (`s9`) with a selectivity block for krill. The inclusion of selectivity blocks aims to, on the one hand, reduce the effect of autocorrelation in the model fitting process, and on the other hand, explicitly account for fleet behavior in exploring new extraction areas, whether in bathymetric or latitudinal terms. According to the data identified, a shift from traditional to continuous trawling occurred in 2006. As a result, two blocks have been established (1980-2006 and 2007-2020) as improvements. This is compared with the base scenario.

@@

\(R\) represents the variability of deviations, adding to the variance caused by environmental factors. \(SpBio\) represent spawning biomass, \(envdata\) is vector of time series of environmental variable. Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual \(R\) should be reduced. However, the model does not automatically adjust for this.

Furthermore, this integrated model took into account environmental variables, as chlorophyll (Chl), known to influence krill abundance and reproductive outputs. These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics. By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.

@@

This stock assessment approach consider predator-prey interactions as a key driver of krill population dynamics. By incorporating information on predator  abundance and feeding ecology, our model provide a more comprehensive understanding of the tropic interactions shaping krill dynamics.  In this analysis we incorporated penguin population index as predator *"fleets"* as we describe previously. This overall natural morality (M) to include explicit mortality (M2) caused by each major predator as M = M1 + sum(M2) [@methot2020stock], and in this way transfer to total mortality in form Z = M1 + M2+ F.

@@

A comparative analysis of models `s01`, `s1`, and `s7` will be conducted, as they incorporate elements such as aggregate data, spatially structured data, and different sets of life history parameters according to @Mardones2024par. Additionally, to address the observations, scenarios `s8` and `s9` have been added and compared. The details of each observation made in the @CCAMLRWGSAM2024 report have been thoroughly addressed, and each response can be reviewed in **APPENDIX 1**.

@@

All models tested was the final gradient using maximum likelihood is achieve and was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis. About goodness-of-fit, Figure \@ref(fig:res1) and Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years in `s1`. In the first graph, we observe multiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS). The second graph shows surveys (SURVEYEI, SURVEYGS, SURVEYJOIN) and predator data (PREDATOR) for `s1`. For many fleets and surveys, greater consistency in krill sizes is observed after the year 2000. However, there was variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity. For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices.

```{r res1,  out.width='60%', fig.cap = "Pearson residuals, comparing across fleets. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected)."}
knitr::include_graphics('s1/plots/comp_lenfit__page1_multi-fleet_comparison.png')
```

```{r res1a, out.width='60%', fig.cap = "Pearson residuals, comparing across fleets. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) (continued)."}
knitr::include_graphics('s1/plots/comp_lenfit__page2_multi-fleet_comparison.png')
```

Regarding this pearson residual analysis, a visual inspection can lead to a mistaken conclusion about the model *goodness-of-fit*. We have integrated a quantitative element to evaluate this type of test. The plotting options are kept mainly to those provided by `r4ss`. With `SSplotRunstest()`, runs tests results are illustrated for mean lengths of size composition data from the  `SS3`model. Green shading indicates no evidence (p > 0.05) and red shading evidence (p < 0.05) to reject the hypothesis of a randomly distributed time-series of residuals, respectively. The shaded (green/red) area spans three residual standard deviations to either side from zero, and the red points outside of the shading violate the ‘three-sigma limit’ for that series. We plot this in Figure \@ref(fig:runtest) and Table \@ref(tab:runtable).

```{r runtest, out.width='40%', fig.show='hold',  fig.cap="Runs test plot and Joint residual plot for mean lengths from fits length composition data from different fleets in Antarctic krill."}
sspar(mfrow = c(5, 2), plot.cex = 0.8)
knitr::include_graphics(c('Figs/run1.png',
                          'Figs/run2.png',
                          'Figs/run3.png',
                          'Figs/run4.png',
                          'Figs/run5.png',
                          'Figs/run6.png',
                          'Figs/run7.png',
                          'Figs/run8.png',
                          'Figs/run9.png'))
```

@@

Regarding the fitting of indices in general, the model attempts to interpret the trends but encounters some difficulties in effectively fitting all sources of information, as shown in (Figure \@ref(fig:resall)).

@@

## Comparision outputs between scenarios

>>>>

---
title: "Integrated approach to modeling krill population dynamics in the Western Antarctic Peninsula: spatial and ecosystem considerations"
author:
  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base]
  - George Watters^[Ecosystem Science Division, Southwest Fisheries Science Center, NOAA, USA]
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
always_allow_html: true
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

@@

The integration of available sources of information is a challenge for any attempt to know dynamics populations for Antarctic Krill (*Euphausia superba*) in an integrated stock assessment model. Being able to identify sources of information, systematize and integrate it in a model and estimate krill population variables in a context of change and fisheries management is vital for management and decision making. The analysis represents a modern approach for stock assessment of complex dynamics such as krill and the integration of ecosystem components as environmental variables and ecological aspects such as predators. We have three main objectives. i)integrate fishing, environmental and ecological variables in an integrated stock assessment model and test his performance; ii) to consider the spatial heterogeneity of the krill population structure and, iii) to evaluate the impact of biological and  population structure assumptions on the performance of this type of integrated stock assessment models. Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management strategies in this region. This study addresses key recommendations provided by the Working Group on Statistics, Assessment, and Modelling (WG-SAM-2024/26) . The base model (s1) performed well in capturing the inter-annual and inter-fleet variability in krill length compositions. Retrospective analysis showed a slight underestimation bias in spawning biomass; however, this was within an acceptable range. The model suggests a higher dependence on length composition data compared to other sources for estimating recruitment. Spatial considerations and assumptions about stock-recruitment relationships significantly impact population variable estimates. The inclusion of a selectivity block into the model (`s9`) improved the fit to some length composition data compared to the base model (`s1`). Overall, this study demonstrates the potential of using a spatially explicit integrated model for krill stock assessment. The model provides valuable insights into krill population dynamics and can be used to inform sustainable management practices.

*Keywords: Krill populations, dynamic population, integrated model, Stock Synthesis (SS3), spatial implicit, ecosystem-based management, CCAMLR*

@@

Antarctic krill (*Euphausia superba*, thereafter krill) population is predominantly concentrated in the Western Antarctic Peninsula (WAP), serving as a vital food source for predators and being targeted by increasing fishing activities. WAP is considered one of the most sensitive areas to Climate Change (CC) and has already experienced fast changes in various dimensions [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a]. Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities. As part of its mandate, CCAMLR promotes sustainable management of krill through an ecosystem-based approach  [@Watters2013]. CCAMLR has tried to conduct different modeling approaches to assess the status of important species, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016]. Various stock assessment approaches on krill have been conducted to provide management advice to CCAMLR using model-based methodologies. Many of these models face challenges in aligning with management schemes or accurately representing the population dynamics of krill, due to its variability and heterogeneity in spatial distribution and ecosystem complexity. Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys [@Green2023; @Watters2013]. However, these approaches may overlook critical aspects of krill dynamics, such as integrated different data components, spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions. There is a growing consensus on the need for using integrated stock assessment models that incorporate multiple sources of information, capture spatial heterogeneity and consider the complex interplay of ecosystem variables in krill. Integrating the complexity of krill population dynamics presents several challenges. Firstly, data consistency over the years has been problematic, as many monitoring programs and studies have not adhered to a uniform standard throughout their temporal and spatial extension. Additionally, certain regions or areas lack sufficient data or possess only limited information regarding krill dynamics. Furthermore, interpreting indirect indicators that may affect the krill population, such as environmental variables, predator abundance, and fishery data, adds another layer of complexity. Despite these gaps, the existing data about krill population dynamics is abundant, considered data rich [@Kawaguchi2024], and this availability allows building an integrated model to get population signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others. In response to this need, we propose a conceptual model to conduct a statistical model in Antarctic krill population in the Antarctic Peninsula, which in turn allows the application of an integrated model that incorporates the fishery, surveys and ecosystem information on krill population over the past 40 years.

@@

The study area for the assessment of krill populations in Subarea 48.1 and strata subdivision to account for spatial heterogeneity [@Dornam2021] (Figure \@ref(fig:mapa)). The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill. Therefore, dividing the study area into strata allows to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.

```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Bransfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
knitr::include_graphics('Figs/481.png')
```

@@

The analysis includes multiple sources of data to inform the population dynamics model listed in Table \@ref(tab:datasources). These data sources include: 1) Haul-by-haul data informed to CCAMLR in the C1 form, which include information about vessel, nationality, and georeferenced catch and effort, particularly useful for CPUE estimates. 2) Data from the CCAMLR Scheme of International Scientific Observation (SISO) program, where records of krill length composition, is systematically collected on board by scientific observers. 3) Data collected from U.S. Antarctic Marine Living Resources (AMLR) research cruises and monitoring programs specifically designed to assess krill abundance by strata, and biological characteristics of krill collected through net sampling. All this data was processed in @Dornam2021; and length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl. Mat. 2; 3).

For this analysis, chlorophyll-a concentration (Chl) was assessed as a primary environmental driver. Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U. Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148). This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and  krill dynamics in the WAP. We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + Sea Ice Concentration (SIC) + Sea Surface Temperature (SST)). The model included spatial component (cell) as random effect (formula: ~1 | cell). Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020] (Suppl. Mat 3; 4). Predator data was analyzed by using penguin abundances at different colonies along WAP available in the Mapping Application for Penguin Populations and Projected Dynamics "MAPPPD" portal. Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.

@@

Source data and temporal scale used in this stock assessment is shown in Figure \@ref(fig:datas2). Data handling for modeling purposes was elaborated with *tidyverse* library and dependencies [@Wickham2019].

@@

A conceptual model for krill in Subarea 48.1 includes  spatial strata described previously. Despite differences among strata, the krill population is closed for emigration and immigration, and their spatial distribution and interactions occur within this single, cohesive population. Ecological and environmental variables that have an impact on the krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels. The model considers interannual fluctuations of growth dynamics , maturity and reproduction, with a spawning pulse per year.

@@

Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables. The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013; @methot2020stock] with the updated version 3.30.21. SS3 is a widely used software tool for assessing fish and invertebrate populations widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions. SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data. The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012] with estimation parameters. The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in the WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that influence the analyzed population. All analyses were executed in R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019]  and *ss3diags* packages [@Winker2023]. Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit).

```{r parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}

@@

parbio<-ctl1$MG_parms[1:10,c(1:4,7)]
 row.names( parbio)<-c("Nat M",
                       "Lmin", 
                       "Lmax",
                       "VonBert K",
                       "CV young",
                       "CV old", 
                       "Wt a", 
                       "Wt b",
                       "L50%", 
                       "Mat slope")
 SRpar<-ctl1$SR_parms[1:5,c(1:4,7)]
 Qpar<-ctl1$Q_parms[1:2,c(1:4,7)]
 Selpar<-ctl1$size_selex_parms[1:22,c(1:4,7)]
 parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))

@@

Like most invertebrates, krill is a species for which aging is complex [@Punt2013]. Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure. Regarding this, we consider a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data in krill monitoring. Aging techniques for krill have not been available historically, although some recent progress has been made [@Kilada2017]. In this sense, reading hard parts is difficult like in krill, catch-at-length data are more plentiful, as the collection of length information is relatively cheap. Length offers insights into the age structure of the population, as there is a correlation between age and length. So far, all integrated models for krill have been of the Catch-at-Length type [@Kinzey2015a; @Kinzey2019a; @Wang2021]. This means they are approaches that internally model age through growth parameters and the Age-length-key (ALK). To perform this conversion, various methods have been used, but the AKL matrix is perhaps the most important one [@ICCA2003]. ALK are generated by aging a sub-sample of a population and used to estimate the age distribution from larger length samples. ALKs assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases. Seasonal or multi-annual applications require careful justification, as using an ALK from a single period can introduce significant bias. Length-stratified sampling is essential, and the process of developing ALKs is labor-intensive, requiring optimal data collection to ensure accuracy  [@ICCA2003]. Formulae exist to estimate the number of age determinations and length measurements is calculated as;

@@

```{r AKL, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in SUbarea 48.1"}
knitr::include_graphics('Figs/AKL.png')
```

@@

Since age data monitoring data was not available, we explored *platoon* (Figure \@ref(fig:platoon)) as an alternative to compare its performance and results with the current length-based model currently used in WG-SAM-2024/26 [@Mardones2023].

```{r platoon, out.width='70%', fig.cap="Representation of platoons in krill length frequencies"}
x_values <- seq(0, 7, by=0.1)
means <- c( 2.9,  3.5, 4.1)
std_devs <- c(0.65, 0.8, 0.65)
data <- data.frame(
  x = rep(x_values, each = length(means)),
  mean = rep(means, times = length(x_values)),
  std_dev = rep(std_devs, times = length(x_values))
)

@@

Here we have incorporated a scenario (`s9`) with a selectivity block for krill. The inclusion of selectivity blocks aims to reduce the effect of autocorrelation in the model fitting process, and also, to explicitly account for fleet behavior in exploring new extraction areas, whether in bathymetric or latitudinal terms. According to the data identified, a shift from traditional to continuous trawling occurred in 2006. As a result, two blocks have been established (1980-2006 and 2007-2020) as improvements. This is compared with the base scenario.

@@

\(R\)  represents the variability of deviations, adding to the variance caused by environmental factors. \(SpBio\) represent spawning biomass, \(envdata\) is a vector of time series of environmental variables. Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual \(R\) should be reduced. However, the model does not automatically adjust for this.

Furthermore, this integrated model took into account environmental variables, such as chlorophyll (Chl), known to influence krill abundance and reproductive outputs. These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics. By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.

@@

This stock assessment approach considers predator-prey interactions as a key driver of krill population dynamics. By incorporating information on predator abundance and feeding ecology, our model provides a more comprehensive understanding of the trophic interactions shaping krill dynamics. In this analysis we incorporated a penguin population index as predator *"fleets"* as we described previously. This overall natural morality (M) to include explicit mortality (M2) caused by each major predator as M = M1 + sum(M2) [@methot2020stock], and in this way transfer to total mortality in form Z = M1 + M2+ F.

@@

A comparative analysis of models `s01`, `s1`, and `s7` will be conducted, as they incorporate elements such as aggregate data, spatially structured data, and different sets of life history parameters according to @Mardones2024par. Additionally, to address the observations, scenarios `s8` and `s9` have been added and compared. To evaluate consistency of the source of information used in each of the tested scenarios, we estimate of R0 through likelihood components.

@@

All models tested were the final gradient using maximum likelihood is achieved and was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis. Regarding goodness-of-fit, Figure \@ref(fig:res1) display the residuals of krill length caught by fleets over the years in `s1`. ultiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS) are recorded Figure \@ref(fig:res1). SURVEYEI, SURVEYGS, SURVEYJOIN and PREDATOR have consistency after the year 2000. However, there was variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity. For instance, in FISHERYBS and FISHERYEI, the distribution widened and slightly shifted towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices.

```{r res1,  out.width='100%', fig.cap = "Pearson residuals, comparing across fleets. red are positive residuals (observed > expected) and blue are negative residuals (observed < expected)."}
knitr::include_graphics('Figs/pearson_heatmap.png')
```

Regarding this Pearson residual analysis, a visual inspection can lead to a mistaken conclusion about the model *goodness-of-fit*. We have integrated a quantitative element to evaluate this type of test. We run a statistical test to illustrate the residual of mean lengths of size composition data from the SS3 model. Green shading indicates no evidence (p > 0.05) and red shading evidence (p < 0.05) to reject the hypothesis of a randomly distributed time-series of residuals, respectively. The shaded (green/red) area spans three residual standard deviations to either side from zero, and the red points outside of the shading violate the ‘three-sigma limit’ for that series (see Figure \@ref(fig:runtest) and Table \@ref(tab:runtable)).

```{r runtest, out.width='40%', fig.show='hold',  fig.cap="Test plot and Joint residual plot for mean lengths from fits length composition data from different fleets in krill."}
sspar(mfrow = c(5, 2), plot.cex = 0.8)
knitr::include_graphics(c('Figs/run1.png',
                          'Figs/run2.png',
                          'Figs/run3.png',
                          'Figs/run4.png',
                          'Figs/run5.png',
                          'Figs/run6.png',
                          'Figs/run7.png',
                          'Figs/run8.png',
                          'Figs/run9.png'))
```

@@

In general,  in the fitting of indices the model attempts to interpret the trends but with some difficulties in effectively fitting all sources of information, in particular, in predator fleet (Figure \@ref(fig:resall)).

@@

## Outputs comparision between scenarios
